# Packages

```{r, message = F}
library(tidyverse)
library(gganimate) # For animation
library(ggthemes)  # For ggplot2 theme_map()
library(magick)    # For combining gifs
```

# Read TidyTuesday data

```{r}
tuesdata <- tidytuesdayR::tt_load(2020, week = 44)
```

# Select key data

Read in key data and rename variables
```{r}
# read in data
data_all <- tuesdata$`wind-turbine` %>% 
#rename columns
  select(id               = objectid, 
         project_name     = project_name,
         year             = commissioning_date,
         capacity_project = total_project_capacity_mw,
         capacity_turbine = turbine_rated_capacity_k_w,
         lat              = latitude,
         lon              = longitude) %>% 
# assign name to Newfoundland project which is missing name 
  mutate(project_name = ifelse(
    id == '1451', 'Newfoundland Project', project_name)) %>% 
# if commissioning date spans multiple years, take the first year
  mutate(year = parse_number(year)) %>%
# arrange by year
  arrange(year, id)
```

# Summarise data by project

Assign each project:
- a location, based on the average of its turbine coordinates;
- overall capacity (MW), based on final capacity achieved; and
- start date, based on year first commissioned.

```{r, message = F}
data_project <- 
  data_all %>% 
  group_by(project_name) %>% 
  summarise(lat = mean(lat),
            lon = mean(lon),
            capacity_project = max(capacity_project),
            year = min(year)) %>% 
  ungroup() %>%
  select(year, everything()) %>% 
  arrange(year, project_name) 
```

# Animated plots

First plot is a line chart showing cumulative capacity by year

```{r}

a <- data_project %>% 
  # calculate cumulative capacity by year
  group_by(year) %>% 
  summarise(capacity = sum(capacity_project)) %>% 
  ungroup() %>% 
  mutate(capacity_cum = cumsum(capacity)) %>% 
  # produce plot
  ggplot(aes(x = year, y = capacity_cum)) +
  geom_line(size = 1, color = 'purple', alpha = 0.5) + 
  geom_point(size = 2, color = 'purple') + 
  transition_reveal(along = year, range = c(1992, 2020))+
  # Add dynamic title that rounds the year to the closest integer
  labs(title = "Canadian wind turbine capacity:  {round(frame_along, 0)}",
       x = "Year", 
       y = "Capacity (MW)")  +
  # other styling
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size=12))

# produce gif
a_gif <- animate(a,  duration = 25, start_pause = 10, end_pause = 15, 
                  height = 400, width = 400)

# save gif
anim_save("turbines1.gif")


```

Second plot is a map showing cumulative capacity for each project by location

```{r}
b <- data_project %>% 
  ggplot(aes(x = lon, y = lat, group = project_name, size = capacity_project)) +
  # Add Canadian borders with a light grey background
  borders(regions = "Canada",
          fill = "grey90",
          colour = "grey90"
          ) +
  geom_point(colour = 'purple', alpha = .2) +
  # range parameter sets the size of each point
  scale_size_continuous(range = c(1, 12), 
                        breaks = c(100, 200, 300)) +
  # Set map projection and theme
  coord_map() +
  theme_map() +
  labs(size = 'Capacity (MW)') +
  # specify the "year" variable as being used in the animation
  transition_reveal(along = year, range = c(1992, 2020)) +
  # add styling
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = c(0.8, 0.6),
        plot.margin=unit(c(1,1,1,0.5),"cm"))

# produce gif
b_gif <- animate(b, duration = 25, start_pause = 10, end_pause = 15, 
                 height = 400, width = 400) 

# save gif
anim_save("turbines2.gif")
```

# Combine plots

Combines the two plots above into the final gif.

```{r}
a_mgif <- image_read(a_gif)
b_mgif <- image_read(b_gif)

new_gif <- image_append(c(a_mgif[1], b_mgif[1]))
for(i in 2:100){
  combined <- image_append(c(a_mgif[i], b_mgif[i]))
  new_gif <- c(new_gif, combined)
}

# save final gif to file
image_write(new_gif, "turbines3.gif")
```

